---
title: "Palmetto and Binomial Regression"
author: "Nicole Schmidt"
date: "2/2/2021"
output:
  html_document:
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, messages = FALSE)

# Attach packages and read in data

library(tidyverse)
library(janitor)
library(here)
library(broom)
library(GGally)
library(jtools)
library(dplyr)
library(kableExtra)


# Read in data and tidy; rename plant species, select variables.

palmetto <- read_csv("palmetto.csv") %>% 
  select(year, plant, species, height, length, width, green_lvs) %>% 
 mutate(species = case_when(species %in% c("1") ~ "Serenoa repens",
                            species %in% c("2") ~ "Sabal etonia"))

palmetto$species <- as.factor(palmetto$species)

```

### Introduction
The palmetto dataset contains "survival and growth data across multiple years, habitats, and experimental treatments" with two dominant palmetto species, _Serenoa repens_ and _Sabal etonia_ at the Archbold Biological Station located in Florida.  

A section containing 2 - 3 finalized data visualizations (with figure captions) in which you explore differences in height, canopy length, canopy width, and green leaves for the two species. Add a figure caption to each. Below each data visualization, add a short (~1 sentence) statement with a takeaway from the visualization.

```{r}

## Visualize the data and look at trends

ggplot(data = palmetto, aes(x = length, y = width))+
  geom_jitter(aes(color=species))+
  facet_wrap(~species)

ggplot(data = palmetto, aes(x = height, y = width))+
  geom_jitter(aes(color=species))+
  facet_wrap(~species)

ggplot(data = palmetto, aes(x = height, y = width))+
  geom_jitter(aes(color=species))+
  facet_wrap(~species)

ggplot(data = palmetto, aes(x = height, y = green_lvs))+
  geom_jitter(aes(color=species))+
  facet_wrap(~species)

ggplot(data = palmetto, aes(x = width, y = green_lvs))+
  geom_jitter(aes(color=species))+
  facet_wrap(~species)

#Observations from data exploration:
#Serenoa repens seems to have more green leaves than Sabal etonia
# Width, length, and height for both plants seem to be positively correlated

```

```{r}
# Finalize Graphs

ggplot(data = palmetto, aes(x = height, y = green_lvs))+
  geom_jitter(aes(color=species))+
  facet_wrap(~species)+
  labs( x = "Height(cm)", y = "Green leaves count", 
        title = "Green leaves count and height comparison between Sabal etonia and Serenoa repens",
        subtitle = ("1981 - 2017"),
        fill = "Species")+
  theme_bw()+
  scale_color_manual(values = c("green", "dark green"))
  

```
<br>
- _Serenoa repens_ has a higher green leaf count than _Sabal etonia_
<br>


```{r}

# Second Graph Height and Width 

ggplot(data = palmetto, aes(x = length, y = width))+
  geom_jitter(aes(color=species))+
  facet_wrap(~species)+
    labs( x = "Length (cm)", y = "Width (cm)", 
        title = "Width and length comparison between Sabal etonia and Serenoa repens",
        subtitle = ("1981 - 2017"),
        fill = "Species")+
  theme_bw()+
    scale_color_manual(values = c("green", "dark green"))

```
<br>
- Both _Sabal etonia_ and _Serenoa repens a positive correlation with length and width. As the length of plants grow so do the width of both plants 
<br>


Binary Logistic Regression 
- length
- width
- height
- green leaves count 
 
```{r}

palmetto_blr <- glm(species ~ length + 
                              width + green_lvs + height, 
                            data = palmetto,
                            family = "binomial")


#checking levels(palmetto$species):
#0 is Serenoa
#1 is Sabel

#when the model is above 50% probability it is serenoa classification
#when the model is below 50% it is likely sabal. 

```

** Table 1.** Binomial Logistic Regression Outcomes for _Serenoa repens_ and _Sabal etonia_
```{r}
# Table 

palmetto <- palmetto_blr %>% 
  broom::tidy() %>% 
  select(estimate, std.error, p.value) %>% 
  mutate_if(is.numeric,funs(as.character(signif(.,3))))

rownames(palmetto) <- c("Intercept",
                        "Length (cm)",
                        "Width (cm)",
                        "Green Leaves (n)",
                        "Height (cm)")

kableExtra::kable(palmetto,
                  col.names= c("Coefficient",
                            "Standard Error",
                            "p-value")) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            position ="center") %>% 
  kableExtra::add_header_above(bold = TRUE,
                               line = TRUE,
                               c("Binomial Logistic Regression for _Serenoa repens_, and _Sabal etonia_" = 4))


```

### Part 3

**Table 2.** Incidents of correct and incorrect predictions of palmetto species based on binary logistic model of predictor variables
```{r}

#model fitting 

blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response") %>% 
  mutate(predicted = ifelse(.fitted >= .5, "Serenoa repens", "Sabal etonia")) %>% 
  mutate(correct = ifelse(predicted == species, "predicted_correctly", "predicted_incorrectly"))
  
palmetto_table <- blr_fitted %>% 
  count(species, correct) %>% 
pivot_wider(names_from = correct, values_from = n) %>% 
mutate(total = rowSums(across(where(is.numeric)))) %>%
mutate(percent = (predicted_correctly / total)* 100) %>%
 mutate_if(is.numeric,funs(as.character(signif(.,3))))


kableExtra::kable(palmetto_table,
                  col.names= c("Species", "Correct Prediction",
                               "Incorrect Prediction", "Total", "% Correct")) %>% 
    kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            position ="center") %>% 
  kableExtra::add_header_above(bold = TRUE,
                               line = TRUE,
                               c("Percent species predicted correctly by binary logistic regression model" = 5))


```


### Data Source
Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5
